---
title: "Parameterized Report (demo)"
author: "Jane Smith (Jacobs)"
date: "`r Sys.Date()`"
params:
  file_plots: NULL
  file_tables: NULL
output:
  officedown::rdocx_document:
    base_format: "bookdown::word_document2"
    reference_docx: !expr here::here('StyleTemplates', 'WordStyleTemplates', 'WordTemplate.docx')
    reference_num: FALSE
    number_sections: FALSE
bibliography: "`r here::here('References', 'TestReferences.bib')`"
link-citations: yes
csl: "`r here::here('References', 'ReferenceStyles', 'chicago-author-date.csl')`"
---

```{r setup, include = FALSE}
# Set knitr options
knitr::opts_chunk$set(echo = FALSE,
                      warnings = FALSE,
                      fig.align = "center",
                      fig.width = 6, fig.height = 4)
# Set other options
## This ensures that Office charts produced by the mschart package match the
## width and height of the figures in the .docx report.
formals(ph_location)[c("width", "height")] <- knitr::opts_chunk$get()[
  c("fig.width", "fig.height")
  ]

# Packages
if (!require("pacman")) install.packages("pacman") # p_load() from pacman checks
# if packages are installed, installs them if not, then loads them.
pacman::p_load(devtools, BiocManager, # for loading/installing other packages
               here, # for using project-relative paths
               cowplot, gridGraphics, # for plotting
               flextable, # for tables
               knitr, bookdown, # for additional knitting options
               officedown, officer, mschart, # for producing Office docs
               tidyverse) # for general data manipulation and piping

# Shiny outputs
## Plots
report_plots <- readRDS(here::here("Temp", params$file_plots))
if(!is.null(report_plots)){
df.plots <- data.frame(item.name = names(report_plots),
                       item.type = "plot",
                       removed = FALSE)
}
## Tables
report_tables <- readRDS(here::here("Temp", params$file_tables))
if(!is.null(report_tables)){
df.tables <- data.frame(item.name = names(report_tables),
                       item.type = "table",
                       removed = FALSE)
}
# Combine
df.items <- rbind(df.plots, df.tables)

# Functions
table.unpack <- function(x){
  purrr::map2_chr(
  .x = names(x),
  .y = x, 
  function(caption, dat) {
    knitr::knit_child(input = here::here("Functions", "child-loop.Rmd"), envir = environment(), quiet = TRUE)}
  ) %>%
  # New page for each table in "appendix"
  cat(sep = "\n\n\\pagebreak\n")
}
```

# Table of contents

<!---BLOCK_TOC--->

## List of figures

<!---BLOCK_TOC{seq_id: 'Fig'}--->

## List of tables

<!---BLOCK_TOC{seq_id: 'Tab'}--->

\newpage

# Section

This report received `r length(report_plots)` plots and
`r length(report_tables)` tables from Shiny.

The names of the plots are: `r names(report_plots)`.

The tables are `r names(report_plots)`.

This was made in R [@R-base].

This is the first plot in the plots output list from Shiny.

```{r fig.cap = "Figure with a caption", fig.id = "Fig1"}
report_plots[[1]]
#item.add(report_plots, item.position = 1)
```

This is the first table in the tables output list from Shiny.

```{r results='asis', echo=FALSE, tab.id = "Tab1", tab.cap = "Table with a caption"}
#item.add(report_tables, item.position = 1) 
report_tables[[1]] %>% head(n = 10) %>% flextable() %>% set_table_properties(layout = "autofit")
```

# References

<div id="refs"></div>

# Appendices

By default, un-rendered plots and tables go into the appendix.

## Plots

Here I print all plots to simulate an appendix.
This includes the plot that was printed, before.

I need to figure out how to identify these as individual plots and pass them to
the table of plots.

```{r, fig.id = "Fig2"}
walk(report_plots, print)
```

```{r}
block_section(prop_section(page_size(orient = "portrait"),
              type = "continuous"))
```

## Tables

Here I print all tables in the output list.

In this case, individual captions will have to be passed to them in the loop,
rather than using the chunk options. I haven't implemented that yet, but those
captions, as well as names, should be editable in Shiny prior to rendering.

```{r results='asis', echo = FALSE}
#item.add(report_tables) %>%
report_tables %>%
  table.unpack()

block_section(prop_section(page_size(orient = "landscape"),
              type = "continuous"))
```